---
title: "Homework 1"
author: "Krista Noe"
date: "February 10, 2021"
output: html_document
---

Use linear regression to determine how beach areas changes through time. The variables are:  
• BeachID: categorical variable, unique identification for each beach  
• Year: continuous variable, year the beach area was measured  
• OpenBeach: continuous variable, the area in hectares of open beach  

***
1. Import the “beach.csv” dataset into R and examine the first few rows of data

```{r}

beach <- read.csv("beach.csv")
head(beach, 3)

```

<br>

2. Fit a model that includes interactions between Year and BeachID. Report estimates of all coefficients

```{r}
fit <- lm(OpenBeach ~ Year * BeachID, data = beach) #so you are predicting area of open beach (response variable), by both year and beach ID (predictor variables)

fit

```

<br>

3. Predict the beach area at all three beaches in 1992

```{r}

# interactive model: y = B0 + (B1*X1) + (B2*X2) + (B3*X1*X2)
# can be rewritten as: y = B0 + (B1+B3)*X1 + B2

a <- 36945.73 + (- 18.36 * 1992)
  # this equation can be written as y = B0 + (B1 * X1) since the other terms equal 0  
  # B0 = 36945.73
  # B1 = -18.36
  # X1 = 1992
  
b <- 36945.73 + (-18.36 + 17.04)*1992 + -34211.98
  # y = B0 + (B1+B3)*X1 + B2
  # B0 = 36945.73
  # B1 = -18.36 
  # B3 = 17.04
  # X1 = 1992
  # B2 = -34211.98


c <- 36945.73 + (-18.36 + 14.37)*1992 + -28776.11
  # y = B0 + (B1+B3)*X1 + B2
  # B0 = 36945.73
  # B1 = -18.36 
  # B3 = 14.37
  # X1 = 1992
  # B2 = --28776.11


print(a)
print(b)
print(c)
```

<br>

4. Plot the predicted area of each beach across all years

```{r}
x = 1992:2014

a <- 36945.73 + (- 18.36 * x)
b <- 36945.73 + (-18.36 + 17.04)*x + -34211.98
c <- 36945.73 + (-18.36 + 14.37)*x + -28776.11

plot(a, xlab = "Year", ylab = "Open Beach Area", type = "o", pch = 17, col = "sienna", xaxt = "n")
title(main = "Beach a")
axis(side = 1, at = 1:23, lab = c("1992", "1993", "1994", "1995", "1996", "1997", "1998", 
                                  "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009",  "2010", "2011", "2012", "2013", "2014"))
plot(b, xlab = "Year", ylab = "Open Beach Area", type = "o", pch = 21, col = "forestgreen", xaxt = "n")
title(main = "Beach b")
axis(side = 1, at = 1:23, lab = c("1992", "1993", "1994", "1995", "1996", "1997", "1998", 
                                  "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009",  "2010", "2011", "2012", "2013", "2014"))
plot(c, xlab = "Year", ylab = "Open Beach Area", type = "o", pch = 8, col = "midnightblue", xaxt = "n")
title(main = "Beach c")
axis(side = 1, at = 1:23, lab = c("1992", "1993", "1994", "1995", "1996", "1997", "1998", 
                                  "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009",  "2010", "2011", "2012", "2013", "2014"))


## now put it all together to compare ##

par(las = 1)
plot(a, type = "o", pch = 17, col = "sienna",
  ylim = c(0, 400), xlab = "Time", ylab = "Beach Open Area", xaxt = "n") 
lines(b, type = "o", pch = 21, col = "forestgreen")
lines(c, type = "o", pch = 8, col = "midnightblue")
title(main = "Beach Open Area Through the Years")
axis(side = 1, at = 1:23, lab = c("1992", "1993", "1994", "1995", "1996", "1997", "1998", 
                                  "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009",  "2010", "2011", "2012", "2013", "2014"))
legend("topright", c("a", "b", "c"), 
      col = c("sienna", "forestgreen", "midnightblue"), 
      lty = c(1, 1, 1), pch = c(17, 21, 8), 
      title = "Beach ID")

```


<br>

5. Interpret the rate of beach area change at each of the 3 beaches 
```{r}

# for beach b and c, open beach area changes by (B1 + B3) hectares
# for beach a, open beach area changes by B1 hectares 

rate_a <- -18.36
rate_b <- (-18.36 + 17.04)
rate_c <- (-18.36 + 14.37)

print(rate_a)
print(rate_b)
print(rate_c)
```
Open area of beach a decreases by 18.36 hectares each year, and has the highest rate of change of all three beaches. In fact, according to the model, by the time data stopped being collected (2014), beach a had 0 hectares of open beach. This does not match the actual recorded data (21.1 hectares), but is not far off. Beach b experiences a decrease of 1.32 hectares per year, and beach c experiences a decrease of 3.99 hectares per year. All three beaches experience a decrease in open area of beach, meaning that there will be less habitat for piping plovers over time on the lakeshores of North Dakota. 


***

Please hand in both the html document and the .Rmd document to GitHub.