---
title: "Homework 3"
author: "Krista Noe"
date: "2/28/2021"
output: html_document
---

- y: response variable (1 for success, 0 for failure)  
- x1: continuous predictor variable  
- x2: categorical predictor variable with 2 levels: "a" and "b"

<br>

1. Fit a logistic regression model that assumes the probability of success is an additive function of variables
x1 and x2.
```{r}
data <- read.csv("Homework 3 Data.csv")
head(data)

fit <- glm(y ~ x1 + x2, family = binomial, data = data)
summary(fit)
```

<br>

2. Interpret the effect of variable x1 on the log odds of success. Verify your interpretation in R.

```{r}
betas <- coef(fit)
betas[1]; betas[2]; betas[3] ##(just to double check they're right!)

p_1 <- plogis(betas[1] + betas[2] * 1 + betas[3] * 0)
p_2 <- plogis(betas[1] + betas[2] * 2 + betas[3] * 0)
# the slope coefficient is negatively correlated; there is a lower probability of ocurrence when x1 is higher 

# log odds ratio
log((p_2 / (1 - p_2)) / (p_1 / (1 - p_1)))

# does it check out in R?
betas[2] 
# yes it does
```


<br>

3. Interpret the effect of variable x2 on the log odds of success. Verify your interpretation in R.

```{r}

p_1 <- plogis(betas[1] + betas[2] * 0 + betas[3] * 0)
p_2 <- plogis(betas[1] + betas[2] * 0 + betas[3] * 1)

log((p_2 / (1 - p_2)) / (p_1 / (1 - p_1)))
betas[3]

# the slope coefficient for x2 is negatively correlated as well; there is a lower probability of ocurrence when x2 is higher 
```


<br>

4. Duplicate the Wald Test and p-values for variables x1 and x2 performed by the glm() function. Do
you reject or fail to reject your null hypothesis?

```{r}
#wald test
ts1 <- betas[2] / summary(fit)[['coefficients']]['x1', 'Std. Error']
ts2<- betas[3] / summary(fit)[['coefficients']]['x2b', 'Std. Error']

#p-value
2 * pnorm(-1 * abs(ts1), mean = 0, sd = 1)  # x1
# fail to reject given that p > 0.05
2 * pnorm(-1 * abs(ts2), mean = 0, sd = 1)  # x2
# reject given that p < 0.05


#checks out with summary output as well
summary(fit)[['coefficients']]['x1', 'Pr(>|z|)']  
summary(fit)[['coefficients']]['x2b', 'Pr(>|z|)']
```



<br>

5. Predict and plot the mean probability of success over the range of values of x1.
```{r}
# log odds of use

#range of x1 values
x1p <- seq(from = min(data$x1), to = max(data$x1), length.out = 100)

#predictions
y <- betas[1] + betas[2] * x1p + betas[3]

plot(x = x1p, y = plogis(y),
xlab = 'x1', cex.axis = 1.5, cex.lab = 1.5, type = 'l')
```
