---
title: "Homework 5"
author: "Krista Noe"
date: "4/4/2021"
output: html_document
---

1. Load data and place into an unmarkedFrameOccu object

```{r}
y <- read.csv("Bobcat.csv")  # detection / nondetection data
p_covs <- read.csv("p covariates.csv")  # detection covariates
site_covs <- read.csv("psi covariates.csv")  # site coviariates

library(unmarked)

bobcat_mat <- as.matrix(y)

det_covs <- list(
  people = data.frame(p_covs[, c(1:71)])
)

occu_data <- unmarkedFrameOccu(y = bobcat_mat, 
                               obsCovs = det_covs, 
                               siteCovs = site_covs)

```

<br>


2. Fit an occupancy model that assumes conditional detection probability is a function of the number of
people photographed at a site, and occupancy probability is a function of disturbance within 5km.

```{r}

fit <- occu(formula = ~ people ~ Dist_5km, data = occu_data)
summary(fit)

```

<br>


3. Interpret the effect of the number of people photographed at each site on conditional detection probability
of bobcat. Verity your interpretation in R.

```{r}
betas <- coef(fit)

p_0 <- plogis(betas[1] + betas[2] + betas[3] + betas[4] * 0)
p_1 <- plogis(betas[1] + betas[2] + betas[3] + betas[4] * 1)

log((p_1 / (1-p_0)) / (p_0 / (1-p_1)))
betas[4]
```
There is a -0.0608 change in the log odds of detecting a bobcat per one person. 


<br> 


4. Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you
reject or fail to reject this null hypothesis?

```{r}
summary(fit)
w <- -23.64 / 4.773
w

2 * pnorm(-1 * abs(w))

```
The null hypothesis for this model is the effect on the log odds success of disturbance within 5km is 0. We reject this null hypothesis because our value (7.313065e-07) is less than 0.05.


<br> 


5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over
the range of observed disturbance within 5km.
```{r}

new_psi <- data.frame(people = rep(0, length.out = 100), 
                      Dist_5km = seq(from = 0, to = 0.13, length.out = 100))

predict <- predict(object = fit, newdata = new_psi,
            type = 'state')

plot(y = predict$Predicted, 
     x = seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100), 
     xlab = "Disturbance Distance", ylab = "Occupancy Probability", type = "l")
lines(y = predict$lower, 
      x = seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100), lty = 2)
lines(y = predict$upper, 
      x = seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100), lty = 2)


```


